\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\geometry{letterpaper, margin=1in}
\usepackage{parskip}
\usepackage{mathpazo}
\usepackage{titling}
\usepackage{cite}  

\setlength{\droptitle}{-0.5in}

\parindent=20pt

\title{Statement of Purpose}
\author{Ray Tsai}
\date{}

\begin{document}


\maketitle

\vspace{-0.25in}

I am applying to the Master's program in Computer Science at Stanford University, specializing in
theoretical computer science to translate my mathematical foundation, particularly in combinatorics,
to abilities that can address real-world problems. My primary interests lie in combinatorial
optimization and cryptography, where the interplay between theory and application is most apparent.
Although I don't have a specific industry goal in mind, I aspire to develop innovative algorithms
and solutions to address challenges in areas like data privacy and resource optimization.

As a mathematics-computer science major at UC San Diego, I placed most of my attention to
mathematics in my undergraduate studies, primarily through honors and graduate-level coursework.
While I appreciate the beauty of pure mathematics, I wasn't satisfied pursuing abstract mathematics
for its own sake. I instead looked for a field in mathematics with a closer connection to the real
world and turned to graph theory, the study of relationships that can model a wide range of
practical problems effectively.

During my sophomore year, I began my research journey by directly reading under Professor Jacques
Verstraete, exploring the vast literature on extremal graph theory. Through the readings, I was
exposed to a variety of powerful techniques for tackling extremal graph problems, such as
probabilistic methods, stability, and finite geometric constructions, which laid the groundwork for
my honors thesis centered on the Double Turán problem. The problem asks for the maximum possible
number of edges across $n$ subgraphs of a complete graph $K_n$ such that no pairwise intersection of
these subgraphs contains a specified forbidden structure. After dedicating my summer to studying the
triangle-free case, I established a tight upper bound on the number of edges under the stricter
condition that each subgraph is induced. Specifically, I showed that the extremal condition is
uniquely achieved when all subgraphs are complete bipartite graphs, by recursively expanding the
intersection of all subgraphs. This case serves as a stepping stone for the project, and I am
currently working on the general triangle-free case. 

Expanding my interests to include the computational aspects of combinatorial problems, I had the
opportunity to learn about computational complexity theory directly from Professor Russell
Impagliazzo, studying Arora and Barak's text, Computational Complexity: A Modern Approach
\cite{arora2009computational}. With this foundation, I then joined Professor Impagliazzo's research
group, studying Multicalibration to mitigate unintended biases to certain subpopulations in learning
models from the perspective of complexity theory. This research opened my eyes to the surprising
connections between abstract mathematical concepts and the practical applications of machine
learning. Specifically, the project brought the application of extremal combinatorics and complexity
theory to a new level by modeling the fairness of algorithms with random-like structures yielded
by Szemeredi's Regularity Lemma, further strengthening my commitment to applying theoretical results
to practical problems.

Stanford's Master's program offers the ideal path to translate my mathematical expertise into a
career focused on developing practical algorithms and solutions. I aspire to contribute to fields
where combinatorial tools are applied extensively, and taking courses like Matching Theory (MS\&E
319), Combinatorial Optimization (CS 261), Algebraic Error Correcting Codes (CS 250), and Advanced
Topics in Cryptography (CS 355) would provide me with the necessary background of combinatorial
applications to computer science that I lack in my undergraduate studies.

Moreover, I am looking forward to the research opportunities at Stanford and aim to pursue
distinction in research. Drawing from my experience in combinatorics, I am most curious to
investigate approximation algorithms. Classic combinatorial optimization problems, such as finding
Hamiltonian cycles and matchings in graphs, frequently arise in real-world contexts, where practical
solutions often rely on approximation algorithms. Thus I hope to utilize my specialized knowledge in
combinatorics to advance the Pareto frontier of these algorithms, pushing the boundaries of their
efficiency and accuracy. Professor Aviad Rubinstein's expertise in approximation algorithms and
algorithmic game theory aligns with my research interest. I am especially intrigued by his algorithm
for approximating maximum matching size in \cite{doi:10.1137/1.9781611977554.ch151}, which beats the
$\frac{1}{2}$-approximation barrier without even requiring a full traversal of the graph! Given the
ubiquity of combinatorial algorithms, I am excited to explore the possibilities of improving these
algorithms under his guidance, making them more accessible for practical applications. If possible,
I would also like to collaborate with Itai Ashlagi from the Management Science and Engineering
department. He has contributed significantly to the area of market design and matching markets, and
I am particularly drawn to his work on kidney exchange. He previously used integer programming to
find long chains in kidney exchange that maximize kidney
transplants\cite{doi:10.1073/pnas.1421853112}, and I am curious if approximate algorithms can be
used to tackle other NP-hard problems regarding this urgent topic.

Additionally, I am interested in studying both the theoretical and practical aspects of cryptography,
particularly in the area of pseudorandomness. Professor Li-Yang Tan's work on the tightness of
Impagliazzo's hardcore theorem \cite{blanc2024samplecomplexitysmoothboosting} revealed the
limitations in hardness amplification for pseudorandom generators (PRG), and I want to study its
implications on the security and efficiency of PRGs. On the other hand, I would also like to join
Professor Dan Boneh's applied cryptography group and work on projects related to pseudorandom
functions.

I have only begun to scratch the surface of theoretical computer science, and there are still a lot
of topics to explore. I am excited about the prospect of joining Stanford's Master's in Computer
Science program to expose me to a broader range of topics in computer science through rigorous
coursework and extensive research opportunities. With the mathematical foundation I built during my
undergraduate studies, I am confident in my ability to excel in the program and evolve into a
capable problem-solver. I look forward to contributing to the vibrant academic community at Stanford
and bridging the worlds of mathematics and computer science.

\newpage

\bibliographystyle{plain}
\bibliography{references}
\end{document}
